---
title: "Project 2"
author: "Kali, Blake, Daniel, and Tony"
date: "12/2/20"
output:
  html_document:
    data_print: paged
---

## Introduction

    For our second project, we decided to use the same data set that we used in our first project which contains key data points collected from various matches in the game League of Legends. League of Legends is a multiplayer online Battle Arena (MOBA) in which two teams of five players each compete to destroy the opposing team’s base, called the Nexus.

    The nexus of each team spawns small, NPC characters called minions (sometimes creeps) that follow one of three lanes and support the player(s) that has chosen that lane to capture objectives and move closer towards the enemy nexus. Each lane is guarded by three towers, each at different locations. These towers also support the player by defeating minions and enemy champions who happen to be in the tower’s range. Once the three towers have been destroyed, the team can move on to destroy the enemy’s inhibitor. This inhibitor prevents the opposing team from summoning an upgraded version of these minions, who deal increased damage and have significantly more health. Each team has three inhibitors, each one corresponding to a lane. Only one needs to be destroyed before the next stage can be completed. The final stage consists of two towers protecting the nexus. Both towers and at least one inhibitor must be destroyed in order to begin damaging the Nexus. The team who destroys their opponent’s Nexus first wins the game.


    The teams are divided into sub roles that players can choose, which dictates the lane that their champion plays in. Each player must choose a champion character to control for the course of the game, and each champion has unique abilities and play styles to fill the roles that each team must have. Some champions excel full frontal encounters, while others are more suited towards stealth and quick strikes, while few can heal and protect their team. Players kill both enemy minions, elite neutral monsters, and enemy players to increase the power of their champion and buy items from the gold that these kills produce. Items increase the stats of the player and allow for the player to “build” into a sort of play style that is suited for the champion they chose.

    In this project, we will attempt to determine what variables in this set are best at predicting a win in League of Legends.

## Import + Cleaning
```{r}
library(tidyverse)
data <- read.csv("https://raw.githubusercontent.com/kalihale/miscfiles/main/high_diamond_ranked_10min.csv")
names(data)[names(data) == "blueWins"] <- "Wins"
data$Wins <- as.factor(data$Wins)
levels(data$Wins) <- c("red", "blue")
data$blueFirstBlood <- as.factor(data$blueFirstBlood)
levels(data$blueFirstBlood) <- c("FALSE", "TRUE")
data$redFirstBlood <- as.factor(data$redFirstBlood)
levels(data$redFirstBlood) <- c("FALSE", "TRUE")

dataBlueWins <- data[data$Wins == "blue", ]
dataRedWins <- data[data$Wins == "red", ]
```

## Data Dictionary

    This data set is a collection of many of the metrics listed above, and will be used in our analysis to see which of these metrics seems to contribute the most to a victory for a given team.For the sake of our analysis, we have set up a dictionary to make our data more understandable. The following is a list of terms and how they are to be understood within the Data Set.

* “Red” – The Color of the Opposing Team

* “Blue” – The Color of the Current Team

* gameId – The Riot ID assigned to the match

* Champion – Refers to the playable character that the player has chosen to compete with.

* Ward – Device used to grant players vision of areas that they are not currently in range to see.

* First Blood - Refers to the first player kill of the game.

* Kill – Refers to the act of a player defeating another player through combat.

* Death – Refers to the act of being defeated in both Player vs. Player and environmental combat.

* Assists – Refers to score points given to teammates who help another player in killing an enemy.

* Lane – Refers to one of the three standard areas in which players compete (Top, Middle, Bottom). They dictate where champions are best suited as well as where minions can be killed.

* Jungle - Refers to the area outside of the lanes where additional monsters exist that can be slain, including large and elite monsters.

* Minions / Creeps – Small non-player characters who roam the lanes of the game, attempting to help players reach the goal of destroying the enemy Nexus.

* Monster – Small non-player characters who reside in the Jungle and are used by champions to gather resources and experience.

* Elite / Elite Monsters – Large,non-player characters who are substantially tougher than the regular monsters and tend to provide some bonus to the player after being slain.

* Dragon – Large monster that resides in the bottom half of the jungle,that comes in four variants. Each variant corresponds to an element, being Earth, Fire, Water, and Air. Dragons provide a bonus depending on the element when killed. When a team kill four dragons, they receive a large bonus.

* Herald – Large monster that resides in the top half of the jungle, when killed drops and eye which can be used by the player to summon a friendly version of the herald that will attack enemy towers and deal massive amounts of damage.

* Baron / Baron Nashor – Large Elite Monster that resides in the top half of the jungle. Typically require a full team effort to kill. When defeated, Baron drops a buff to the team which killed him that increases the effectiveness that minions have.

* Gold - Currency collected in the game used to buy equipment for champions.

* Experience - Points used to gauge progress towards champion levels, awarded by killing players, monsters, and minions.

* Level – Refers to the current tier that the chosen champion currently resides as. Each level comes with increased value of that champion.

* CS / Creep Score / MS / Minion Score – Score that each player has which is a sum tally of how many non-playable enemies that the player has killed.

* Diff – Refers to the difference in between two measurements.


# Data Analysis

    In this section, we are going to explore if the elite monsters on the map will affect the winrate. Since the dragon spawn at 5:00 and respawn every 5 minutes after killed, the Herald spawns at 8:00 and respawns once 6 minutes after being killed, there could only be 1 being capture for the first 10 minutes of a game.
```{r}
#Make data$Wins numerical to calculate the winrate
levels(data$Wins) <- c("0", "1")
data$Wins <- as.numeric(data$Wins)
##Select Data
Blue_dragon <- data[data$blueDragons==1,]
Red_dragon <- data[data$redDragons==1,]

Blue_Herald <- data[data$blueHeralds==1,]
Red_Herald <- data[data$redHeralds==1,]

Blue_both <- data[data$blueEliteMonsters==2,]
Red_both <- data[data$redEliteMonsters==2,]

nrow(Blue_dragon)/nrow(data)
nrow(Red_dragon)/nrow(data)
##Red side have slightly more chance to get a dragon
nrow(Blue_Herald)/nrow(data)
nrow(Red_Herald)/nrow(data)
##Blue side have slightly more chance to get a herald
nrow(Blue_both)/nrow(data)
nrow(Red_both)/nrow(data)
##About the same

```
### Does dragon make a difference in winrate and gold?

    H0: No. The mean difference of winrate is less than or equals to 0.
    Ha: Yes. The mean difference of winrate is greater than 0.

```{r}
data$blueDragons <- as.factor(data$blueDragons)
data$redDragons <- as.factor(data$redDragons)
data$blueHeralds <- as.factor(data$blueHeralds)
data$redHeralds <- as.factor(data$redHeralds)
ggplot(data = data, aes(x = blueDragons, y = blueTotalGold)) + geom_violin()
t.test(data$Wins,Blue_dragon$Wins)$p.value
```

    Very small number. We can reject the null hypothesis. We can conclude that getting dragon can definitly increase the winrate and, from our plot, it appears to also increase the gold of the team that gets it.

### Does Herald make a difference in winrate and gold?

    H0: No. The mean difference of winrate is less than or equals to 0.
    Ha: Yes. The mean difference of winrate is greater than 0.
```{r}
ggplot(data = data, aes(x = blueHeralds, y = blueTotalGold)) + geom_violin()
t.test(data$Wins,Blue_Herald$Wins)$p.value

```

    Very small number. We can reject the null hypothesis. We can conclude that getting herald can definitly increase the winrate and, from our plot, it appears to also increase the gold of the team that gets it.

### How about both?

    H0: No. The mean difference of winrate is less than or equals to 0.
    Ha: Yes. The mean difference of winrate is greater than 0.
```{r}
t.test(data$Wins,Blue_both$Wins)$p.value
```

    Very small number. We can reject the null hypothesis. We can conclude that getting both dragon and herald can definitly increase the winrate.

```{r, echo = FALSE}

data$Wins <- as.factor(data$Wins)
levels(data$Wins) <- c("red", "blue")

```
## Experience diff

    The experience difference a team has is positively related to the their performance in game. If a team has a positive experience diff in comparison to the other team (current - opposing), they exhibit greater frequency of wins.
```{r}
ggplot(data = data, aes(x = Wins, y = blueExperienceDiff, fill = Wins)) + geom_violin()
t.test(dataBlueWins$blueTotalExperience, dataBlueWins$redTotalExperience)
t.test(dataRedWins$blueTotalExperience, dataRedWins$redTotalExperience)

```

## Gold diff


    The creep score of a team vs its gold gain rate per minute are positively related. The more minions a team kills, the quicker and more gold their tend to gain.
```{r}
ggplot(data = data, aes(x = blueTotalMinionsKilled, y = blueGoldPerMin, color = Wins)) + geom_violin()
```


## Wards Placed
```{R}

meanPerGame <- function(default = 1){
  
  return( (data$blueWardsPlaced[default] + data$redWardsPlaced[default]) )
  
}
means <- sapply(1:nrow(data), meanPerGame)

mean(means)
```
```{R}

max(dataBlueWins$blueWardsPlaced)
max(dataBlueWins$redWardsPlaced)
max(dataRedWins$redWardsPlaced)
max(dataRedWins$blueWardsPlaced)

```

    Plots of the Distribution of Wards based upon wins-- despite winning or losing, neither team shows any inherent advantage based upon their wards. 
```{R}
ggplot(data = data, aes(x = Wins, y = blueWardsPlaced, fill = Wins)) + geom_violin()
ggplot(data = data, aes(x = Wins, y = redWardsPlaced, fill = Wins)) + geom_violin()

length(which(data$Wins == "blue"))
length(which(data$Wins == "red"))

ggplot(data = data, aes(blueWardsPlaced, redWardsPlaced, color = Wins)) + geom_count() + geom_smooth()
ggplot(data = data, aes(x = Wins, y = blueWardsDestroyed, fill = Wins)) + geom_violin()
ggplot(data = data, aes(x = Wins, y = redWardsDestroyed, fill = Wins)) + geom_violin()
ggplot(data = data, aes(blueWardsDestroyed, redWardsDestroyed, color = Wins)) + geom_point() + geom_smooth()

```

    Boostrap assumptions & t Test for the mean of blue wins, all three seem to show little correlation between performance improvement and placing & destroying wards.
```{R}
x <- function(default=1){
  resample <- sample(dataBlueWins$blueWardsPlaced, length(dataBlueWins), replace = T)
  return(mean(resample))
}
functionData <- sapply(1:10000, x)

sum(functionData > mean(dataRedWins$blueWardsPlaced)) / length(data)

y <- function(default=1){
  resample <- sample(dataBlueWins$blueWardsDestroyed, length(dataBlueWins), replace = T)
  return(mean(resample))
}
functionData <- sapply(1:10000, y)
sum(functionData > mean(dataRedWins$blueWardsDestroyed)) / length(data)


t.test(dataBlueWins$blueWardsPlaced,dataRedWins$blueWardsPlaced)$p.value
```
## Kills & First Blood
  
    Based upon these violin / box plots, we can assume an inverse relationship between red deaths and red performance in game. The more deaths red has, the less their frequency of winning. 
```{r}

ggplot(data = data, aes(x = blueFirstBlood, y = blueTotalGold)) + geom_violin() + geom_boxplot(width = 0.5)


t.test(data$blueTotalGold, as.numeric(data$blueFirstBlood))$p.value
```

    Deaths appear to have a more significant effect on the outcome of the game. The same assumption applies to blue.

```{r}

ggplot(data = data, aes(x = Wins, y = blueDeaths, fill = Wins)) + geom_violin() + geom_boxplot(width = 0.5)

```

```{R}
ggplot(data = data, aes(x = Wins, y = blueKills, fill = Wins)) + geom_violin()
```

# Models and Analyze
Setup Train and Test
```{r}
set.seed(12345)
data$ID <- 1:nrow(data)
train <- sample_frac(data,0.6)
test <- anti_join(data,train,by="ID")
train$ID <- NULL
test$ID <- NULL
```

## k Nearest Neighbors
```{r}
knnTrain <- train
knnTest <- test
library(class)
train_Wins <- knnTrain$Wins
test_Wins <- knnTest$Wins
knnTrain$Wins <- NULL
knnTest$Wins <- NULL
knnTrain$blueFirstBlood <- NULL
knnTest$blueFirstBlood <- NULL
knnTrain$redFirstBlood <- NULL
knnTest$redFirstBlood <- NULL
```

```{r}
predicted_Wins <-knn(knnTrain,knnTest,train_Wins,k=1,prob=TRUE)

table(predicted_Wins,test_Wins)
```
K nearest neighbors seems to get the result a little over half right. Unfortunately, if we increase k, the model only gets less accurate.

## Logistic Regression
Logistic Regression does not particularly help us. There isn't a significant division between any groups from the observations in the data set, so we omitted the model from use.


## Decision Trees
```{R}
library(tree)

model <- tree(Wins ~ ., data = train)
prune.misclass(model, best = 2)
summary(model)
plot(model)
text(model, pretty = 0)



predictedWins <- predict(model, test, type ="class")
actualWins <- test$Wins

table(predictedWins, actualWins)

```
This model gives us about a 73% accuracy, which is interesting considering exactly one decision is made on this decision tree. The gold difference between teams would have a large effect on games, since gold is used to buy items which can give players an advantage, and the larger the gold difference is, the bigger the advantage of the team which is ahead.

##  Random Forests

```{R}
library(randomForest)

train$ID <- NULL
test$ID <- NULL
rTree <- randomForest(Wins ~ ., data = train, ntree = 200, mtry = 2, importance = TRUE)
varImpPlot(rTree, n.var= 10, sort = T, scale = T)

predictedWins <- predict(rTree, test)
actualWins <- test$Wins


table(predictedWins, actualWins)
```

    Despite this data originating from the first 10 minutes of gameplay in high-tier games, the random forest does extremely well in its accuracy. Holding approximately a 75% accuracy rating, the random forest model does well in predicting the team which ultimately wins the game. We used this model because it typically tends to perform at the very least on par with if not much better than the other models we typically use, and not only this, we get a much more in-depth look at the factors which the model relies most on. For instance, the gold and experience differences between the two teams seemed to contribute most to the prediction, which makes sense in game. The team with the most gold can afford to buy the required equipment for their champions faster and then outperform the opposing team, which leads to an increase in the rate of experience.


# Conclusions & Further Questions

    From our analysis, we can conclude that gold and experience are the two largest factors in determining the winner of a game. Since gold and experience are directly related to what goes on in a game and largely related to other variables that we analyze here, we tried running a version of our models with all of the gold and experience information set to NULL, and to our surprise, the models actually significantly decreased in accuracy. This indicates that experience and gold are, to some extent, independent from the other variables in this dataset. This is most likely due to the fact that there are ways to gain experience and gold that are not directly related to other variables, such as through the use of items that boost experience when used or double the amount of gold going to a team. Unfortunately, this dataset does not give us any information about what items the team bought. An analysis on a dataset which includes that information might indicate whether the type and stats of an item changes the chances of a given team winning.
    The random forest model gave us the most accurate prediction, with the decision tree close behind. K nearest numbers came in a distant third with a little over 50% accuracy, and the linear regression model did not work at all for this set of data.
    These models would be interesting and useful to gamers who play League of Legends and teams hoping to improve their strategy. Players would be able to focus on a strategy that increases their experience and gold, therefore increasing their chances of winnning.
